---
title: "lab4_hw"
format: html
author: "Megan Hessel"
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| output: false 
library(LexisNexisTools)
library(quanteda)
library(tm)
library(topicmodels)
library(topicdoc)   
library(tidyverse)
library(tidytext)
library(reshape2)
library(tictoc)
library(LDAvis)
library(tsne)
library(servr)
```

## Load in Data

```{r}
#| output: false 
#tbl <- read_csv(here::here("week5", "Nexis", "ocean-acidification"))

pre_files <- list.files(
  path = here::here("week5", "Nexis", "ocean-acidification"),   # update to your folder
  pattern = "\\.docx$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

pre_dat <- lnt_read(pre_files)

pre_meta_df       <- pre_dat@meta
pre_articles_df   <- pre_dat@articles
pre_paragraphs_df <- pre_dat@paragraphs

df <- tibble(
  date     = pre_meta_df$Date,
  headline = pre_meta_df$Headline,
  id       = pre_meta_df$ID,
  text     = pre_articles_df$Article,
  source   = pre_meta_df$Newspaper
)

```

# 1. Build and clean my corpus

## Build Corpus

```{r}
#| warning: false 
#| output: false 
tbl <- df
corpus <- corpus(x = tbl, 
                 textg_field = "text")
```

## Tokenize and Clean

```{r}
# Stop words 
add_stops <- c(quanteda::stopwords("en", 
                                   source="smart"),
               # remove other meta data words 
               "said", "says", "told", "according", "discuss", "because", "where", "vol", "pg", "newstex")

# Tokenize 
toks <- tokens(corpus, 
               remove_punct = TRUE, 
               remove_numbers = TRUE, 
               remove_symbols = TRUE) %>% 
  # Normalize
  tokens_tolower() 

# Remove stop words 
toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")
```

## Build the Document-Feature Matrix

```{r}
# Creating a dfms
dfm1 <- dfm(toks1) 
dfm2 <- dfm_trim(dfm1, min_docfreq = 2) # Reduce by limiting min values of feature's doc frequency

# Remove empty rows b/c LDA will error on these
sel_idx <- slam::row_sums(dfm2) > 0 # select no empty rows 
dfm <- dfm2[sel_idx,] # subset
```

# 2. Select K using multiple critera

## Step 1: Initial Model (Theory-Driven k)

Run 3 models (K = 3)

```{r}
# Define Hyperparameter 
k <- 3

# Model 
topicModel_k3 <- LDA(dfm, # matrix
                      k, 
                      method = "Gibbs", # Gibbs is default (always do this)
                      control = list(iter = 100, # Number of iterations
                                     verbose = 25, # Tell us after every 25 iterations
                                     seed = 321
                                     ))
```

### Inspect the Posterior Distributions

```{r}
result <- posterior(topicModel_k3)

beta  <- result$terms
theta <- result$topcis

cat("phi dimensions  (topics × vocab):", dim(beta), "\n")

# Top 10 terms per topic
terms(topicModel_k3, 10)
```

**K=3 Model Results/Groups:**

There are 3 topics (k=3):

1.  climate, al, access, open, change

2.  ocean, coreal, marine, reefs, corals

3.  coral, reef, fig, species, carbonate

(The first words in each list are the most probable word for each topic
/highest probability of word occurrence)

## Step 2: Selecting k with Coherence Scores

To select the best K, will run the model over a range of Ks. The
coherence scores measure how related the words within a generated topic.
Therefore, the score will give use a good idea on an optimal K.

```{r}
# Fit model for a range of k values
k_values <- seq(2, 15, by = 1) 

tic() # Timerstart

# Running the model 
coherence_results <- map_dfr(k_values, function(k){ # Loop thru Ks  
  # Run the model 
  m <- LDA(dfm, k = k, 
      method = "Gibbs", 
      control = list(iter = 300, 
                     seed = 231, 
                     verbos = 0))
  # Calculate coherence 
  coh <- mean(topic_coherence(m, dfm))
  tibble(k=k, coherence = coh)
  }) 

toc() # Timer end 
```

Now, lets plot the coherence vs K!

```{r}
ggplot(coherence_results, aes(x = k, y = coherence)) +
  geom_line(color = "darkorchid4") +
  geom_point(color = "darkorchid4", size = 2) +
  geom_vline(xintercept = coherence_results$k[which.max(coherence_results$coherence)], # Best coh score 
             linetype = "dashed", color = "violetred2") +
  labs(title = "Mean topic coherence by number of topics (k)",
       subtitle = "Dashed line = k with highest coherence",
       x = "Number of topics (k)",
       y = "Mean coherence score") +
  theme_minimal()
```

**Optimization metrics:** The peak coherence (the "elbow") is at K = 4
with a K of 3 following closely behind. Interestingly, a K of 5 drops
significantly, meaning the region is unstable.

**Theory:** Possible ocean acidification themes: coral reef biology,
ocean chemistry, climate policy/governance, and marine organism impact.

**Interpretability**: With a K =3 (from Step 1), topics 2 and 3 are
practically the same groupings. Due to this heavy overlap, K=3 is
under-splitting meaning a K of 4 could be favorable.

The coherence score shows K=4 being favorable. Four ocean acidification
themes come to mind. Lastly, step 1's initial model with a K of 3 is
under-splitting. Thus, a K of 4 seems to be optimal.

# **3. Fit your final model and visualize top terms**

## Step 3: Fit the Selected Model

```{r}
#| output: false 

# Updated with the peak coherence K (k = 5) 
k <- 4  

topicModel_k_select <- LDA(dfm, k,
                           method = "Gibbs",
                           control = list(iter = 1000, 
                                          verbose = 25,
                                          seed = 321))
```

Show the 4 topics:

```{r}
tmResult <- posterior(topicModel_k_select)
terms(topicModel_k_select, 10)
```

```{r}
# Values per topic per text 
theta <- tmResult$topics
# Values per vocab word 
beta  <- tmResult$terms
# Vocab words 
vocab <- colnames(beta)
```

## Step 4: Visualize Top Terms per Topic

Visualizing the 5 topics and the most probable words for each topic.

```{r}
topics <- tidy(topicModel_k_select, matrix = "beta")

top_terms <- topics |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  arrange(topic, -beta)

top_terms |>
  mutate(term = reorder_within(term, beta, topic, sep = "")) |>
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free_y") +
  scale_x_reordered() +
  coord_flip() +
  labs(title = "Top 10 terms per topic",
       subtitle = "β = P(term | topic)",
       x = NULL, y = "β") +
  theme_minimal()
```

## Step 5: Name the Topics

Naming the topics

Assign a descriptive name to each topic (not just the concatenated top
terms).

```{r}
# Option 1: auto-name from top 5 terms
# topic_words <- terms(topicModel_k_select, 5)
#topic_names <- apply(topic_words, 2, paste, collapse = " ")

#Option 2: Replace topic names with interpretive labels after inspection
topic_names <- c("climate_change", "coral_reef_ecology",
                 "reef_biodiversity",  "research")

topic_names
```

## Step 6: Document-Topic Distributions (θ)

Distribution of topics on Documents

Plot θ for the first 6 document.

```{r}
# Grabbing first 6 documents
example_ids <- c(1:6)
n <- length(example_ids)

example_props <- theta[example_ids, ]
colnames(example_props) <- topic_names

# Value per topic per document 
viz_df <- melt(cbind(data.frame(example_props),
                     document = factor(1:n)),
               variable.name = "topic",
               id.vars = "document")

# Plot
ggplot(data = viz_df, aes(topic, value, fill = document)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_flip() +
  facet_wrap(~ document, ncol = n) +
  labs(title = "Topic distributions across sample documents",
       subtitle = "θ = P(topic | document)",
       x = NULL, y = "Topic proportion")
```

## Step 7: Interactive Visualization with LDAvis

```{r}
#| output: false 

svd_tsne <- function(x) tsne(svd(x)$u)

json <- createJSON(
  phi           = tmResult$terms,
  theta         = tmResult$topics,
  doc.length    = rowSums(dfm),
  vocab         = colnames(dfm),
  term.frequency = colSums(dfm),
  mds.method    = svd_tsne,
  plot.opts     = list(xlab = "", ylab = "")
)

#serVis(json)
```

**LDAvis**: do the topic circles separate well, or do they overlap
heavily?

-   The circles do not overlap at all (which is surprising). Each circle
    is on opposite sides of the graph with no overlap.
-   The topics are quite proportional with each topic holding about
    20-30% of the tokens.

**Interpret your topics:**

Topic 1 focus on broad impacts of climate change on oceans and coastal
environments. Words like "change", "ocean", and "warming" led me to this
conclusion.

Topic 2 focus on coral reef ecosystems, specifically their biological
and chemical dynamics. Words like "carbonate" and "fish" led me to this
conclusion. "Analysis" and "data" suggest a research-oriented approach
to studying coral reef ecosystems.

Topic 3 focus on coral bleaching and its effects on reef biodiversity
and marine life. Words like "bleaching", "coral", and "species" led me
to this conclusion. "Scientists" and "world" suggest a global scientific
concern or international research attention surrounding coral reef
decline.

Topic 4 appear to combine climate change, carbon, energy, and scientific
communication . Words like "research", "change", and "carbon" led me to
this conclusion.

Words that surprised me: fig, data, analysis, al, pdf. These metadata
words should have been taken out at the beginning as they are irrelative
to the thematic content. "Fig" is most likely due to figure captions. I
originally thought "al" was "AI" miss-spelled. However, I think "al"
refers to citations "et al”. The thematically scientific words are not
surprising.
