---
title: "Lab5_hw"
format: html
author: "Megan Hessel"
---

## Set up

Load libraries

```{r packages, include = FALSE}
library(tidytext)
library(tidyverse)
library(tidymodels)
library(textrecipes)
library(naivebayes)
library(discrim) 
library(glmnet)
```

Load data

```{r}
urlfile ="https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/climbing_reports_model_dat.csv"
incidents_df<-readr::read_csv(url(urlfile))
```

20/80 test and training split

```{r}
set.seed(321)

# 2 classes 
incident2class <- incidents_df %>% 
  mutate(deadly = factor(if_else(
    is.na(Deadly), "nondeadly", "deadly"
  )))

# 20/80 split 
incidents_split <- rsample::initial_split(incident2class, prop = 0.8, strata = deadly)
incidents_train <- training(incidents_split)
incidents_test <- testing(incidents_split)

```

## K-Nearest Neighbor

------------------------------------------------------------------------

\<Identify the key hyperparameter(s) for your chosen algorithm. Briefly explain in plain language what the hyperparameter controls and why tuning it matters.\>

I am running K-nearest neighbor for the incidents data set. The key hyper-parameter for KNN is the number of neighbors (K). KNN predicts the classification for a value based on the classifications of the surrounding values. K represents the number of neighbors that determine how values will be classified. If K is too big, the model is underfitting whereas a small K causes overfitting.

For instance, we are trying to predict a home's income in Houston, TX. If we use all of Harris county to predict the home's income (large K), the prediction will be too broad. If we only use the home's direct neighbor to predict the home's income (small K), the prediction will be bad as gentrification can cause one "nice" home to be neighbors to a lesser-nice home. Therefore, it is important to pick the correct K for the data set.

To pick a good K value, I will tune K using cross validation and select the best hyper parameter for the data set. Then I will fit the final tuned model and evaluate it on the test set and access the accuracy and plot ROC and the confusion matrix.

----

### Tuning Hyper-parameter

Tune at least one hyperparameter using cross-validation and select a final value.

```{r}
# Penalty-tuning-specification 
knn_spec <- nearest_neighbor(neighbors = tune(), ) %>% 
  set_engine("kknn") %>% 
  set_mode("classification")

# Create Cross-Validation Folds
cv_folds <- vfold_cv(incidents_train, v = 10, strata = deadly)

# Hyper parameter - K neighbors 
k_grid <- grid_regular(neighbors(range = c(1, 50)), levels = 20)
```

```{r}
# Create a workflow to bundle recipe and model - updated workflow 
knn_wf <- workflow() %>% 
  add_recipe(recipe) %>% 
  add_model(knn_spec)

# TUNE - Fit the model 
set.seed(321)
tune_knn <- tune_grid( # tune_grid() fits a model at each of the HP values 
  knn_wf,
  cv_folds,
  grid = k_grid,
  control = control_resamples(save_pred = T))

# Evaluate & Predictions
chosen_acc <- tune_knn %>% select_by_one_std_err(metric = "accuracy", neighbors)
```

### Fit Tuned Model 

Fit your final tuned model and evaluate it on the test set.

```{r}
# Finalize our workflow with best regularization penalty.
final_knn  <- finalize_workflow(knn_wf, chosen_acc)

# Fit to the training data.
fitted_knn <- fit(final_knn, incidents_train)

```

```{r}
# Fit to the test data 
last_fit(final_knn, incidents_split) %>%
  collect_metrics()

# Get Predictions
predictions <- last_fit(final_knn, incidents_split) %>% 
  collect_predictions()
```

### Plots: ROC and Confusion Matrix

Create an ROC plot

```{r}
# ROC 
predictions %>%
  group_by(id) %>%
  roc_curve(truth = deadly, .pred_deadly) %>%
  autoplot() +
  labs(
    color = "Resamples",
    title = "ROC curve for Climbing Incident Reports"
  )
```

Compare its accuracy and ROC AUC to the Naive Bayes and Lasso models from the demo. Which model performs best, and why do you think that is?

Examine the confusion matrix for your tuned model.

```{r}
# Confusion Matrix 
tune_knn %>%
  conf_mat_resampled(parameters = chosen_acc, # pass the best parameters (AI told me this)
                     tidy = FALSE) %>%
  autoplot(type = "heatmap") + 
  scale_fill_gradient(low = "#f7fbff", high = "#08306b")
```

What does the confusion matrix tell you about how the model is handling the minority class? Is overall accuracy a good summary metric here? What would you prioritize — minimizing false negatives (missed fatalities) or false positives (false alarms) — and why?
