---
title: "hw-Lab2"
output: html_document
author: "Megan Hessel"
---

```{r}
library(patchwork)
```

## **7. Interpretive questions:**

1.  What do the most positive keywords (overrepresented in 2025) tell you about how data center coverage has shifted?

The biggest difference in 2025 compared to 2021 is the high frequency of 'Trump' and 'AI'. In 2025, there is also words like 'tariffs', 'nuclear', 'bannon', 'power', and 'energy'. These all show the shift towards politics and energy politics instead of technical infrastructure. Today, data center coverage is being covered through the AI boom lens as well as seen as an expansion to the political environment. Additionally, there is a common story line displaying data centers demand for energy.

2.  Are any of these keywords surprising? What might explain their appearance?

'altman', 'marcos', 'bannon' and 'doutthat' are surprising to me.

-   Sam Altman is the CEO of Open AI.

-   Marcos might refers to the San Marcos Data Center project. The \$1.5 billion, 200 megawatt project was highly protested within the community and was rejected in the San Marcos City Council in February of 2026. Or it might refer to the Phillippines' Marcos administration as the Phillippines become a regional data center hub. 

-   Steve Bannon was employed as the White House chief strategist under Trump.

-   Douthat is a New York Times opinion columnist.

I am surprised these names occurred enough to flag them. This illustrates that coverage is becoming more personality-driven and political.

3.  Keyness flags statistical unusualness, not substantive importance. A high-G² word could be driven by a single prolific source or a short news spike. How would you check whether a keyword is genuinely distributed across the corpus rather than concentrated in one place?

You can check by investigating overall frequency, frequency by source, and monthly frequency. This checks distribution of words on various levels: source, overtime, and overall, making sure the words are not concentrated in one place. 

4.  How has the language used to describe data centers in news coverage changed over time, and what does that suggest about the evolving journalistic frame around the issue?


In 2021, the language is more localized, technical, and business oriented. In 2025, the language shifts towards more national contentment and energy politics. The language has changed from being about "where and who is building the infrastructure" to "how is the infrastructure impacting political and electrical power".  

But, we have to be careful when using Keyness because it only shows the relative change. There can be similarities between time frames, but they are not shown because they are stable. 

------------------------------------------------------------------------

## 8. Exercises

1.  **Stopword sensitivity:** Remove *all* domain-specific stopwords and re-run the frequency plot. Then add 10 more domain stopwords beyond what was provided. How do your results change? What does this tell you about the analytical choices embedded in preprocessing?

    *Your answer here*
I REALLY DO NOT UNDERSTAND WHAT THEY WANT ME TO DO. HOW DO I PICK THE WORDS TO GET RIDE OF. THIS SEEMS BIAS TO JUST REMOVE WORDS BECAUSE "**I** DONT THINK THEY ARE RELEVENT". 
    ```{r stopword-sensitivity}

  # Remove all domain-specidic stopwords 
  domain_sw <- tibble(word = c(
   # News wire / attribution boilerplate
    "said", "says", "told", "according", "reported", "noted", "added",
    "stated", "announced", "mr", "ms", "inc", "corp", "llc",
    # Generic quantitative terms ##(adjust based on your research question)
    "year", "years", "percent", "million", "billion", "quarter",
    "company", "companies", "firm", "business",
    # Consider whether these are substantive for our RQ:
    "data", "center", "centers" 
  ))
      
  # Remove domain-specidic stopwords 
  tokens_clean <- tokens_clean |>
    anti_join(domain_sw1, by = "word") |>
    filter(
      !str_detect(word, "^[0-9]+$"),    #remove pure digit tokens... No pure numbers 
      str_length(word) > 2            #remove very short tokens
  )
  
  # Re-Run the frequency plot 
  top_words <- tokens_clean |>
    count(word, sort = TRUE) |>
    slice_head(n = 30)
  
  frequency_plot1 <- top_words |>
    ggplot(aes(x = reorder(word, n), y = n)) +
    geom_col(fill = "#1a9641") +
    coord_flip() +
    labs(
      title = "30 Most Frequent Terms — Data Center Corpus",
      subtitle = "After stopword removal and lemmatization",
      x = NULL,
      y = "Frequency"
    ) +
    theme_minimal(base_size = 13)
      
  # ------- 
  
 # Remove all domain-specidic stopwords + 10 more 
  domain_sw <- tibble(word = c(
   # News wire / attribution boilerplate
    "said", "says", "told", "according", "reported", "noted", "added",
    "stated", "announced", "mr", "ms", "inc", "corp", "llc",
    # Generic quantitative terms ##(adjust based on your research question)
    "year", "years", "percent", "million", "billion", "quarter",
    "company", "companies", "firm", "business",
    # Consider whether these are substantive for our RQ:
    "data", "center", "centers" 
    # Add 10 more words 
  ))
      
  # Remove domain-specidic stopwords 
  tokens_clean <- tokens_clean |>
    anti_join(domain_sw1, by = "word") |>
    filter(
      !str_detect(word, "^[0-9]+$"),    #remove pure digit tokens... No pure numbers 
      str_length(word) > 2            #remove very short tokens
  )
  
  # Re-Run the frequency plot 
  top_words <- tokens_clean |>
    count(word, sort = TRUE) |>
    slice_head(n = 30)
  
  frequency_plot2 <- top_words |>
    ggplot(aes(x = reorder(word, n), y = n)) +
    geom_col(fill = "#1a9641") +
    coord_flip() +
    labs(
      title = "30 Most Frequent Terms — Data Center Corpus",
      subtitle = "After stopword removal and lemmatization",
      x = NULL,
      y = "Frequency"
    ) +
    theme_minimal(base_size = 13)
  
  # Compare the 2 plots 
  frequency_plot1 | frequency_plot2
  
    ```

2.  **Alternative keyness comparison:** Instead of comparing time periods, compare the two sources with the most articles. What does keyness reveal about how each publication frames the data center beat?

    *Your answer here*

    ```{r alternative-keyness}

    ```

3.  **Bigrams:** Use `unnest_tokens(token = "ngrams", n = 2)` to extract bigrams. What are the 20 most frequent bigrams? Do they reveal entities or conceptual phrases that single-token analysis misses?

    *Your answer here*

    ```{r bigrams}

    ```

------------------------------------------------------------------------
